Tables

### 2019-07-19 code/Supplement-figures-tables.Rmd
Formatting of table .rmd output to make it equivalent to paper tables

---
title: "Supplement - Biserial correlation transformations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

@@

@@

het %>% knitr::kable(align = "c", booktabs = TRUE, escape = FALSE, digits = 2, col.names = c("RP", "Effect", "K", "Effect type", "Effect size estimate",  "$I^2$(\\%)", "$I^2$ 95\\% CI", "$\\tau$", "$\\tau$ 95\\% CI", "r*" , "r* $\\tau$", "r* $\\tau$ 95\\% CI", "Zero", "Small", "Medium", "Large")) %>% 
  add_header_above(c(" " = 12, "Level of heterogeneity" = 4)) %>% 
  add_header_above(c(" " = 12, "Type I Error Rate & Statistical Power" = 4)) %>% 
  kable_styling("scale_down", font_size = 7) %>% #comment out if outputting html file
  column_spec(5, width = "5em") %>% #make sure the 'Effect size estimate' header is wrapped 
  footnote(threeparttable = TRUE, escape = FALSE,
           general = "Effects were estimated in metafor using REML. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. RP = Replication Project, K = no. primary studies, CI = confidence intervals, r* = effect sizes as correlations and biserial correlations. $I^2$ and its confidence intervals are for original effect size specifications whereas $\tau$ and its confidence intervals are for transformed r* effect sizes. Statistical power and type I error rates were simulated, where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75\\\\%) respectively. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation. Code to reproduce table: osf.io/kf6pt/",
           alphabet = "These effects were simulated as standardized mean differences") %>% 
  landscape()

@@

>>>>

---
title: "Tables and Figures - Supplement A and B"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

This document contains the code to reproduce tables and figures from Supplement A and Supplement B for the project "Heterogeneity in direct replications in psychology and its association with effect size".

@@

##Supplement A

@@

het %>% knitr::kable(align = "c", booktabs = TRUE, escape = FALSE, digits = 2, col.names = c("RP", "Effect", "K", "Effect type", "Effect size estimate",  "$I^2$(\\%)", "$I^2$ 95\\% CI", "$\\tau$", "$\\tau$ 95\\% CI", "r*" , "r* $\\tau$", "r* $\\tau$ 95\\% CI", "Zero", "Small", "Medium", "Large")) %>% 
  add_header_above(c(" " = 12, "Level of heterogeneity" = 4)) %>% 
  add_header_above(c(" " = 12, "Type I Error Rate & Statistical Power" = 4)) %>% 
  kable_styling("scale_down", font_size = 7) %>% #comment out if outputting html file
  column_spec(5, width = "5em") %>% #make sure the 'Effect size estimate' header is wrapped 
  footnote(threeparttable = TRUE, escape = FALSE,
           general = "Effects were estimated in metafor using REML. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. RP = Replication Project, K = no. primary studies, CI = confidence intervals, r* = effect sizes as correlations and biserial correlations. $I^2$ and its confidence intervals are for original effect size specifications whereas $\tau$ and its confidence intervals are for transformed r* effect sizes. Statistical power and type I error rates were simulated, where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75\\\\%) respectively. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation. Code to reproduce table: https://osf.io/j9zyd/?view\\\\_only=e6639d8d4f924739a7310782fbbb4e06",
           alphabet = "These effects were simulated as standardized mean differences") %>% 
  landscape()

@@

##Supplement B

```{r Supplement-B-table}

#Summary of results. These are incorporated into the main table in the paper, see tables.rmd
zero <- dens %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4', Effect = effect)

##Medium effect size results
dens2 <- readRDS("../data/appendixA_power_simulation_results.RDS") 
names(dens2) <-  sort(unique(dat$effect))

#Summary of results.
medium <- dens2 %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4') %>% 
  select(-effect)

##combine for table

comparison <- cbind(zero, medium)
```

Sensitivity analysis assuming a medium true effect size rather than zero true effect size

In our main results we simulated data to estimate type I error and power using an average true effect size equal to zero. As explained in the methods section this is unlikely to matter as our interest lies in heterogeneity. However, as a sensitivity analysis, in Supplement B we also estimated type I error and power, but generated with an average 'medium' true effect size as defined by Cohen (1988).

Estimated type I error and power for zero/small/medium/large heterogeneity when simulated with a true effect size of zero vs. medium are shown for each effect in Table A1. In both cases type I error is approximately nominal, as compared to the expected 5% error rate. With a true effect size of zero, average power was `r 100*round(mean(zero$Small), digits = 2)`%/`r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`% for small/medium/large heterogeneity, respectively. The corresponding power with a true medium effect size was also `r 100*round(mean(zero$Small), digits = 2)`%/ `r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`%, respectively. To conclude, changing true effect size in our simulations from zero to medium had little to no impact on estimated type I error and power.

_Type I error and power to detect four heterogeneity levels under two true effect sizes_

```{r}
knitr::kable(comparison, digits = 2)
```

**Note.** Column headers indicate the degree of simulated heterogeneity where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75%) respectively. Medium effect size is as defined by Cohen (1988). Code to reproduce table: https://osf.io/j9zyd/?view_only=e6639d8d4f924739a7310782fbbb4e06

### 2019-07-19 code/tables.rmd
Formatting of table .rmd output to make it equivalent to paper tables

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
##Table 1
###Effect size $\rho_{xy}$ and its heterogeneity as a function of true effect size and measurement reliability.  
```{r effect-size-heterogeneity-table, cache = TRUE}
#library(kableExtra)
rooted_relabilities <- matrix(rep(c(.6, .7,.8), 3), ncol = 3)
rho <- c(0, .3, .5)
rel <- c("$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .6", "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .7", "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .8")

@@

rooted_relabilities %>% knitr::kable(booktabs = T, col.names = c("", "$\\rho_{xy}$ = 0", "$\\rho_{xy}$ = .3", "$\\rho_{xy}$ = .5"), escape = FALSE, digits = 2) %>% 
  kable_styling(position = "left") %>% 
  footnote(threeparttable = TRUE, escape = FALSE,
           general = "Values in cells are observed effect sizes arising from the true effect size $\\\\rho_{xy}$ and measurement reliabilites $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ when sample size is infinite. Code to reproduce table: osf.io/kf6pt/?view_only=e6639d8d4f924739a7310782fbbb4e06") %>% 
  row_spec(3, hline_after = T)

@@

description <- c("Two correlational effects: 'Gender math attitude' compared implicit attitudes (IAT) towards math between genders and 'IAT correlation math' correlated implicit attitudes with self-reported measures. The remainder were experiments with two independent groups. The groups were primed in some way (Anchoring 1-4; low vs. high category scales; norm of reciprocity; flag priming; currency priming), asked to imagine slightly different situations (Sunk costs; gain vs. loss framing; gambler's fallacy; imagined contact) or asked their agreement with statements presented differently (Allowed vs. forbidden; quote attribution). ", #ML1
            "Several effects were experiments with two independent groups. The groups were either primed in some way (Power and perspective; warmth perceptions; subjective distance interaction), saw slightly different statements (Elaboration likelihood interaction; credentials interaction) or experienced different situations (weight embodiment). Examined interactions were between treatment conditions and participant characteristics. One priming effect (metaphor) compared two treatment groups with a control. One effect was correlational: 'persistence and conscientiousness' was measured by an unsolvable anagram task and self-report respectively. The Stroop task is a within-person experiment with two conditions and the 'Availability' effect asks participants to judge whether some letters are more common in the first or third position.", #ML2
                 "Verbal overshadowing 1; Independent two-group experiment. Participants either described a robber after watching a video or listed countries/capitals and after a filler task attempted to identify the robber in a lineup.", #RRR1
            "Verbal overshadowing 2; Different from 1 only in that the filler task took place before the descriptive task instead of after.", #RRR2
                 "Grammar's effect on interpretation; Independent two-group vignette experiment with three outcome variables. Participants read about actions either described in imperfect or perfect tense and then rated protagonist's intentions (intentionality/intention attribution/detailed processing).", #RRR3
            "Ego depletion; Independent two-group experiment. Participants either assigned to a cognitively demanding task or a neutral, and performance was then measured in a subsequent cognitive task.", #RRR4
            "Commitment on neglect/exit; Independent two-group experiment with two outcome variables. Participants either primed to think about commitment to or independence from partner.", #RRR5
            "Facial feedback hypothesis; Independent two-group experiment. Participants either induced to 'smile' or 'pouth' by holding a pen in their mouth differently and simultaneously rated funniness of cartoons. ", #RRR6
            "Intuitive cooperation; Independent two-group experiment. Economic game with money contribution to a common pool either under time pressure or time delay.",
                 "Professor priming; Independent two-group experiment. Participants primed with either a 'professor' or 'hooligan' stimuli. Outcome was percentage correct trivia answers.") #RRR8

@@

#Combine dataframes to create full table
het <-  het %>% 
  left_join(I2_power) %>% 
  arrange(desc(s_I2),  desc(s_ci.ub)) %>% #sort by I2 and upper CI bound
  mutate(I2_ci = paste0("[", trimws(format(round(s_ci.lb, 1), nsmall = 1)), ", ", trimws(format(round(s_ci.ub, 1), nsmall = 1)), "]"), #Round and put confidence interval between brackets, e.g., [0, 0.41]
         s_I2 = format(round(s_I2, 1), nsmall = 1),
         tau_ci = paste0("[", trimws(format(round(tau_ci.lb, 3), nsmall = 3)), ", ", trimws(format(round(tau_ci.ub, 3), nsmall = 3)), "]"),
         tau = format(round(tau, 3), nsmall = 3)) %>% #rounded here instead of in table function because should have 4 digits
  select(rp, effect, k, effect_type, eff_size, s_I2, I2_ci, tau, tau_ci, zero, small, medium, large)

@@

het %>% knitr::kable(align = "c", booktabs = TRUE, escape = FALSE, digits = 2, col.names = c("RP", "Effect", "K", "Effect type", "Effect size estimate",  "$I^2$(\\%)", "$I^2$ 95\\% CI", "$\\tau$", "$\\tau$ 95\\% CI", "Zero", "Small", "Medium", "Large")) %>% 
  add_header_above(c(" " = 12, "Level of heterogeneity" = 4)) %>% 
  add_header_above(c(" " = 12, "Type I Error Rate & Statistical Power" = 4)) %>% 
  kable_styling("scale_down", font_size = 7) %>% #comment out if outputting html file
  column_spec(5, width = "5em") %>% #make sure the 'Effect size estimate' header is wrapped 
  footnote(threeparttable = TRUE, escape = FALSE,
           general = "Effects were estimated in metafor using REML. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. RP = Replication Project, K = no. primary studies, tau = between studies standard deviation, CI = confidence intervals. Statistical power and type I error rates were simulated, where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75\\\\%) respectively. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation. Code to reproduce table: osf.io/kf6pt/?view_only=e6639d8d4f924739a7310782fbbb4e06",
           alphabet = "These effects were simulated as standardized mean differences") %>% 
  landscape()

>>>>

##Table 1
###Effect size $\rho_{xy}$ and its heterogeneity as a function of true effect size and measurement reliability.  
```{r effect-size-heterogeneity-table, cache = TRUE}
#library(kableExtra)
rooted_relabilities <- matrix(rep(c(.6, .7,.8), 3), ncol = 3)
rho <- c(0, .3, .5)
rel <- c("$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .6", "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .7", "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .8")

@@

rooted_relabilities %>% knitr::kable(booktabs = T, col.names = c("", "$\\rho_{xy}$ = 0", "$\\rho_{xy}$ = .3", "$\\rho_{xy}$ = .5"), escape = FALSE, digits = 2) %>% 
  kable_styling(position = "left") %>% 
  row_spec(3, hline_after = T)

```
*Note:* Values in cells are observed effect sizes arising from the true effect size $\rho_{xy}$ and measurement reliabilites $\sqrt{R_{xx'}}\times\sqrt{R_{yy'}}$ when sample size is infinite. Code to reproduce table: osf.io/kf6pt/?view_only=e6639d8d4f924739a7310782fbbb4e06

@@

description <- c("Two correlational effects: 'Gender math attitude' compared implicit attitudes (IAT) towards math between genders and 'IAT correlation math' correlated implicit attitudes with self-reported measures. The remainder were experiments with two independent groups. The groups were primed in some way (Anchoring 1-4; low vs. high category scales; norm of reciprocity; flag priming; currency priming), asked to imagine slightly different situations (Sunk costs; gain vs. loss framing; gambler's fallacy; imagined contact) or asked their agreement with statements presented differently (Allowed vs. forbidden; quote attribution). ", #ML1
            "Several effects were experiments with two independent groups. The groups were either primed in some way (Power and perspective; warmth perceptions; subjective distance interaction), saw slightly different statements (Elaboration likelihood interaction; credentials interaction) or experienced different situations (weight embodiment). Examined interactions were between treatment conditions and participant characteristics. One priming effect (metaphor) compared two treatment groups with a control. One effect was correlational: 'Conscientiousness and persistence' was measured by an unsolvable anagram task and self-report respectively. The Stroop task is a within-person experiment with two conditions and the 'Availability' effect asks participants to judge whether some letters are more common in the first or third position.", #ML2
                 "Verbal overshadowing 1; Independent two-group experiment. Participants either described a robber after watching a video or listed countries/capitals and after a filler task attempted to identify the robber in a lineup.", #RRR1
            "Verbal overshadowing 2; Different from 1 only in that the filler task took place before the descriptive task instead of after.", #RRR2
                 "Grammar's effect on interpretation; Independent two-group vignette experiment with three outcome variables. Participants read about actions either described in imperfect or perfect tense and then rated protagonist's intentions (intentionality/intention attribution/detailed processing).", #RRR3
            "Ego depletion; Independent two-group experiment. Participants either assigned to a cognitively demanding task or a neutral, and performance was then measured in a subsequent cognitive task.", #RRR4
            "Commitment on neglect/exit; Independent two-group experiment with two outcome variables. Participants either primed to think about commitment to or independence from partner.", #RRR5
            "Facial feedback hypothesis; Independent two-group experiment. Participants either induced to 'smile' or 'pouth' by holding a pen in their mouth differently and simultaneously rated funniness of cartoons. ", #RRR6
            "Intuitive cooperation; Independent two-group experiment. Economic game with money contribution to a common pool either under time pressure or time delay.",
                 "Professor priming; Independent two-group experiment. Participants primed with either a 'professor' or 'hooligan' stimuli. Outcome was percentage correct trivia answers.") #RRR8

@@

#Combine dataframes to create full table
het <-  het %>% 
  left_join(I2_power) %>% 
  arrange(desc(s_I2),  desc(s_ci.ub)) %>% #sort by I2 and upper CI bound
  mutate(I2_ci = paste0("[", trimws(format(round(s_ci.lb, 1), nsmall = 1)), ", ", trimws(format(round(s_ci.ub, 1), nsmall = 1)), "]"), #Round and put confidence interval between brackets, e.g., [0, 0.41]
         s_I2 = format(round(s_I2, 1), nsmall = 1),
         tau_ci = paste0("[", trimws(format(round(tau_ci.lb, 3), nsmall = 3)), ", ", trimws(format(round(tau_ci.ub, 3), nsmall = 3)), "]"),
         tau = format(round(tau, 3), nsmall = 3)) %>% #rounded here instead of in table function because should have 3 digits
  mutate_at(vars(zero, small, medium, large), ~format(round(., 2), nsmall = 2)) %>% #round here because of superscripts added later
  select(rp, effect, k, effect_type, eff_size, s_I2, I2_ci, tau, tau_ci, zero, small, medium, large)

#Add superscripts
het <- het %>% 
  mutate(s_I2 = ifelse(effect %in% c("Grammar on intention attribution", "Conscientiousness and persistence"), paste0(s_I2, "\\textsuperscript{a}"), s_I2), #add a superscript
         tau = ifelse(effect == "Conscientiousness and persistence", paste0(tau, "\\textsuperscript{a}"), tau)) %>% 
  mutate_at(vars(zero, small, medium, large), ~ifelse(effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales', 'Verbal overshadowing 1', 'Verbal overshadowing 2'), paste0(., "\\textsuperscript{b}"), .)) #add b superscript

@@

het %>% knitr::kable(align = "c", booktabs = TRUE, escape = FALSE, digits = 2, col.names = c("RP", "Effect", "K", "Effect type", "Effect size estimate",  "$I^2$(\\%)", "$I^2$ 95\\% CI", "$\\hat\\tau$", "$\\hat\\tau$ 95\\% CI", "Zero", "Small", "Medium", "Large")) %>% 
  add_header_above(c(" " = 9, "Level of heterogeneity" = 4)) %>% 
  add_header_above(c(" " = 9, "Type I Error Rate & Statistical Power" = 4)) %>% 
  kable_styling("scale_down", font_size = 7) %>% #comment out if outputting html file
  column_spec(5, width = "5em") %>% #make sure the 'Effect size estimate' header is wrapped 
  footnote(threeparttable = TRUE, escape = FALSE,
           general = "Effects were estimated in metafor using REML. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. RP = Replication Project, K = no. primary studies, $\\\\hat\\\\tau$ = between studies standard deviation, CI = confidence intervals. Statistical power was simulated, where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large
heterogeneity ($I^2$ = 25/50/75\\\\%) respectively. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation. Code to reproduce table: osf.io/kf6pt/?view\\\\_only=e6639d8d4f924739a7310782fbbb4e06" ,
           alphabet = c("Value rounded to zero", "These effects were simulated as standardized mean differences")) %>% 
  landscape()
